# Variables, Math, and Comparisons

## Help

```{r tidy=FALSE, eval=FALSE}
# Hi. This is a comment.

# If you know a function's name, but not how to use it:
?t.test
```

You can also click a function and press F1 to see the help page.

If you don't know the exact name of a function or variable, you can type part of the name and press tab to autocomplete and see some info about it.

## Assignment

```{r tidy=FALSE}
a = 6
b = 8
c = 5.44
d = TRUE
e = "hello world" 
e = 'hello world' # same as double quote
```

*Note: No semi colon or "var" needed*

You'll sometimes see `a <- 6` instead of `a = 6`. Just use `=`. Some people insist on using `<-`. They are silly.

## Names with weird characters

R allows names to have a `.`, and it's common in many built-in functions. For your own variables, avoid it if possible. If you want to have a space in a name, use an underscore (`_`) instead of being ridiculous.

*To learn how to access object members, see the* [lists chapter](lists.html#accessing-elements-in-a-list)*.*

```{r}
this.is.a.variable.name = 1
better_name = 2
```

You can use any weird character like a space in a variable name by surrounding the name with \`. Avoid it if you can, but sometimes it's necessary when you load data from a file.

```{r}
`more than four (>4)` = 5
```

## Console Output

Print `a` in the console

```{r}
a
```

*The `[1]` is output because it is the first element in an array. For more info, see the [arrays chapter](arrays.html).*

Another option that's useful inside functions, which don't output most results

```{r}
print(a)
```

## Math

Arithmetic

```{r tidy=FALSE}
z = a + b
z = a - b
z = a * b
z = a / b
z = a %/% b # Integer division
z = a %% b # Note the double % for the modulo operator
z = a ^ b # exponent

1 + 2 - 3 * 4 / 5 ^ 6 # Please excuse my dear aunt, Sally
```

*Note: There is no ++ or +=*

Functions for floats

```{r tidy=FALSE}
floor(4.82)
ceiling(4.82)
```

Rounding

```{r}
round(4.4) # round down
round(4.6) # round up
round(4.5) # round to even (down)
round(5.5) # round to even (up)
```

Other basic math functions

```{r tidy=FALSE}
sin(pi/2) + cos(0) # radians, not degrees
log(exp(2)) # base e (like ln) is the default
log(100, 10) # use base 10
```

## Comparisons

```{r}
a == b
a != b
a > b
a < b
a >= b
a <= b
```

## Boolean

```{r}
TRUE & FALSE
TRUE | FALSE
!TRUE
```

There's `&` and `&&`. You usually want just `&`.
