
# Variables, Math, Comparisons, and Strings

## Help 

```{r tidy=FALSE, eval=FALSE}
# Hi. I'm a comment.

# If you know a function's name, but not how to use it:
?chisq.test

# You can also type part of the name and press tab to autocomplete and see a bit info

# if you know what you want to do, but don't know the function name:
??chisquare
```



## Assignment

```{r tidy=FALSE}
a = 6
b = 8
c = 5.44
d = TRUE
e = "hello world" 
e = 'hello world' # same as double quote
```
*Note: No semi colon or "var" needed*

You'll sometimes see `a <- 6` instead of `a = 6`. Just use `=`. Some people insist on using `<-`. They are silly.

## Console Output

Print `a` in the console
```{r}
a
```
*The `[1]` is output because all values are arrays.*

Another option useful inside functions
```{r}
print(a)
```


## Names with weird characters

You can use any weird character like a space in a variable name by surrounding the name with \`. Avoid it if you can, but sometimes it's necessary when you load data from a file.
```{r}
`more than four (>4)` = 5
`more than four (>4)`
```

## Math

Arithmetic
```{r tidy=FALSE}
z = a + b
z = a - b
z = a * b
z = a / b
z = a %/% b # Integer division
z = a %% b # Note the double % for the modulo operator
z = a ^ b # exponent

1 + 2 - 3 * 4 / 5 ^ 6 # Please excuse my dear aunt, Sally
```
*Note: There is no ++ or +=*

Functions for floats
```{r tidy=FALSE}
floor(4.82)
ceiling(4.82)
```

Rounding
```{r}
round(4.4) # round down
round(4.6) # round up
round(4.5) # round to even (down)
round(5.5) # round to even (up)
```

Other basic math functions
```{r tidy=FALSE}
sin(pi/2) + cos(0) # radians, not degrees
log(exp(2)) # base e (like ln) is the default
log(100, 10) # use base 10
```

## Comparisons

```{r}
a == b
a != b
a > b
a < b
a >= b
a <= b
```

## Boolean

```{r}
TRUE & FALSE
TRUE | FALSE
!TRUE
```
There's `&` and `&&`. You usually want just `&`.

## Strings

Strings are not arrays in R, so array techniques may not work on strings.

String length
```{r}
nchar('hello world')
```

Substring
```{r}
substring('hello world', 2, 10)
```

Comparison
```{r tidy=FALSE}
'hello' == "hello"
```

### Strings with special characters

If you want to use special characters in a string, you need to "escape it" by adding `\`
```{r}
"string with backslashes \\, double quote \", and unicode \u263A"
```

Or you can use the literal `r"(text)"` which is useful for a Windows path or regular expression
```{r}
r"(c:\hello\world)"
```



### String Concatenation

Concatenate with a space in between
```{r}
paste('hello', 'world')
```
Use a difference separator
```{r}
paste('hello', 'world', sep='_')
```
No separator
```{r}
paste('hello', 'world', sep='')
```