# Types

For this chapter, you'll need to load the tidyverse library. [More on libraries](libraries-and-packages.html)

R is dramatic and gives "conflict" messages that look like errors. Ignore them.

```{r message=FALSE}
library(tidyverse)
```

## Numbers

R has integers but defaults all numbers to `numeric` which is a double precision float.

```{r}
x = 5 # no decimal but still a double
y = x + 1
```

Good ol' float point comparison

```{r}
x = .58
y = .08
x - y == 0.5
round(x-y, digits=1) == round(0.5, digits=1)
```

Numeric division returns a double

```{r}
9 / 2 # double precision float
9 %/% 2 # drop the part after the decimal
```

## Strings

Single and double quotes are the same in R, but a given string needs the same in the beginning and end

```{r tidy=FALSE}
"hello world"
'hello world'
"single quote ' in a string"
'double quote " in a string'
```

Strings are not character arrays in R, so array techniques may not work as expected.

String length

```{r}
length('hello world') # R sees this is one string, not many charaters
str_length('hello world') # the actual length of the string
```

Substring

```{r}
str_sub('hello world', 2, 10)
```

Comparison

```{r tidy=FALSE}
'hello' == "hello"
```

### Strings with special characters

If you want to use special characters in a string, you need to "escape it" by adding `\`

```{r}
"string with backslashes \\, double quote \", and unicode \u263A"
```

Or you can use the literal `r"(text)"` which is useful for a Windows path or regular expression

```{r}
r"(c:\hello\world)"
```

### String Concatenation

Concatenate with a space in between

```{r}
paste('hello', 'world')
```

Use a different separator

```{r}
paste('hello', 'world', sep='_')
```

No separator

```{r}
paste('hello', 'world', sep='')
paste0('hello', 'world') # a shortcut function for no separator
```

See the [strings chapter in R4DS](https://r4ds.had.co.nz/strings.html) for more.

## Dates

See the [dates chapter in R4DS](https://r4ds.had.co.nz/dates-and-times.html).

## Finding the type of a variable

```{r}
class(c(5, 'hi', TRUE))
```

## Checking the type

What's the type?

```{r}
class(5)
```

Remember, arrays are the same as single values.

```{r}
class(1:5)
```

Test if numeric

```{r}
is.numeric(5)
```

Test if string

```{r}
is.character('hi')
```

Test if boolean

```{r}
is.logical(TRUE)
```

## Converting and parsing

Parse or convert to numeric

```{r}
as.numeric(c("5", TRUE, 1:3, "abc"))
```

To string

```{r}
as.character(5)
format(1/3)
format(1/3 , digits = 16)
as.character(TRUE)
```

Convert to boolean. Zero is false. Other numbers are true.

```{r}
as.logical(0:2)
```

## Special types

### NA

Missing values are very common in datasets.

```{r}
is.na(c(NA, 1, ""))
```

Any operation performed on NA will also yield NA. So, you can operate on arrays with missing values.

```{r}
c(5, NA, 7) + 1
```

Be careful about aggregation functions like `min()`, `max()`, and `mean()`. To ignore NAs, use the `na.rm` parameter.

```{r}
mean(c(5, NA, 7), na.rm=TRUE)
```

### Factor

A factor is like an enum in other languages. It encodes strings as integers via a dictionary.

Create an array with many repeating values

```{r}
data = sample(c("hello", "cruel", "world"), 12, replace=TRUE)
data
```

Make it into a `factor`

```{r}
data = factor(data)
data
```

The array is now an integer array with a dictionary

```{r}
as.numeric(data)
data[1]
```

See the different values in the array

```{r}
levels(data)
```

For more info, see [forcats](https://forcats.tidyverse.org/).
