
# Functions

Basic function
```{r}
foo = function () {
  print("hello world")
}
foo()
```
*Note: in the function, you need to use `print` to output*

## Parameters

Parameters and return values
```{r}
addOne = function (x) {
  return(x + 1)
}
addOne(5)
```
*The syntax for return is like a function: `return(value)`*

Parameter order can be arbitrary
```{r}
add = function (x, y) {
  return(x + (y*10))
}
add(x=2, y=10)
add(y=10, x=2)
```

Functions are vectorized by default
```{r}
addOne(1:5)
```

All parameters are pass-by-value because functions are immutable.

```{r}
a = 5
foo = function (a) {
  a = 6
  print(paste("Inside the function as a parameter: ", a))
}
print(paste("Before the function: ", a))
foo(1)
print(paste("After the function: ", a))

```


## Scope

When you assign a value inside a function, it creates a local variable in the scope of the function. You can't access the global variable inside the function. (OK, you can, but the syntax is so obnoxious that I pretend it doesn't exist)

```{r}
a = 5
foo = function () {
  a = 6
  b = 100
  print(paste("Inside the function a =", a))
  print(paste("Inside the function b =", b))
}
print(paste("Before the function a =", a))
foo()
print(paste("After the function a =", a))
#trying to use `b` will cause an error because it is out of scope
```


## A function in a function

Might be useful for encapsulation
```{r}
foo = function (a, b) {
  square = function(x) {
    return(x ^ x)
  }
  print(c(a, b))
  print(c(square(a), square(b)))
}
foo(1, 10)
```


## Dot dot dot
```{r}
foo = function (a, b) {
  return (a / b)
}
bar = function(a, ...) {
  return(foo(a, ...))
}
bar(50, 10)
bar(b = 10, 50) # named works too
```

## Operators

Operators like `+` or `-` or even `[` are all functions. To use them like a function, surround them with \`.

```{r}
`+`(3, 4)
`*`(3, 4)
`[`(5:10, 2) # you don't need the close bracket (])
```

