
# Lists

A list is like an array but it can multiple types of elements.

## Make a list
```{r tidy=FALSE}
x = list(
  a = 5,
  b = 2,
  Long_Name = 4.8,
  "named with spaces" = 0,
  12, # not every element needs a name
  a = 20 # names don't have to be unique (but you really should avoid this)
)
```


## Accessing elements in a list

Get a tuple of the key and value
```{r}
x['a'] # by key name
x[1]   # by index
```

Multiple keys
```{r}
x[c('b', 'a')] # by key name
x[c('b', 'a')] # by key name
```

Type the list name, then `$`, and press tab. R will pop up a list of keys to autocomplete. R uses `$` in the way that other languages use `.`
```{r}
x$Long_Name
```
*Note: Only the value is returned.*

## Brackets for real

Sometimes, you need to use double brackets. Why? Best I can figure out, R is a snarky shit and doesn't believe you actually want what you said you want. So you need to use double brackets to explain to R that you're sure this is what you want.

Double brackets only works for single items, not subsetting.
```{r}
x[['a']] # by key name
x[[1]]   # by index
```



## Names and values

Use the `names()` function to get and set names. It behaves like an array.
```{r}
names(x)
names(x)[3]
```

You can modify names by assigning strings to the resurned array. This is weird. Take a minute to let it sink in.
```{r}
names(x) = c("first", "second", "third", "fourth")
names(x)[3] = "new name"
```

If all elements are the same type, this will get a vector of values
```{r}
myList = list(a=1, b=2, c=3, d=4)
as.vector(unlist(myList))
```