# Arrays

In R, arrays are commonly called "vectors". R likes to be special.

## Everything is an array

In R, even single values are arrays. That's why you see `[1]` in front of results: even single values are the first item in an array of length one.

## Creation

`c()` is some sort of legacy nonsense from the S language. I think it means *character array* even though it can hold things other than characters.

I pronounce it "CAW". Like the sound a crow makes.

Simple array

```{r}
c(8, 6, 7, 5)
```

For multiple types, R converts elements to the most complex type (usually a string). For a real multi-typed collection, see [lists](lists.html)

```{r}
c(9, 'hello', 7)
```

## Array generators

R has a cultural fear of complete words. Many terms are shortcuts or acronyms.

Repeat

```{r}
rep(0, 4)
rep(c(1,2,3), 4) # repeate the whole array
rep(c(1,2,3), each=4) # repeat each item in the array before moving to the next
```

Sequence

```{r}
#increment by 1
4:10
#increment by any other value
seq(from=10, to=50, by=5)
```

Randomly sample from a given distribution

```{r}
# uniform distribution (not 'run if')
runif(n=5, min=0, max=1)
# normal distribution
rnorm(n=5, mean=0, sd=1)
```

## Combining arrays

An array made up of smaller arrays concatenates them. R doesn't seem to allow for an array of arrays.

```{r}
x = 1:3
y = c(10, 11) # arrays of arrays get flattened
z = 500

c(x, y, z)
```

*Note: `z` is technically an array of length 1*

Collapse an array into a string

```{r}
paste(1:5, collapse=", ")
```

## Indexing

```{r}
a = 10:20
```

Get the first value - **Indices start at 1, not 0**

```{r}
a[1]
```

2nd and 6th values

```{r}
a[c(2,6)]
```

Exclude the 2nd and 6th values

```{r}
a[c(-2,-6)]
```

Range of values

```{r}
a[2:6]
```

Any order or number of repetitions

```{r}
a[c(2, 4, 6, 6, 6)]
```

Specify values to keep or drop using booleans (keep this in mind for the "Array operations" section)

```{r}
a[c(TRUE, FALSE, TRUE, FALSE, TRUE, FALSE, TRUE, FALSE, TRUE, FALSE)]
```

## Sampling from an Array

Randomly sample from an array. Elements may repeat.

```{r}
sample(1:3, size=10, replace=TRUE)
```

*`replace` means "sample with replacement", so an element can be sampled more than once*

Sample without replacement. Elements will not repeat.

```{r}
sample(1:5, size=4, replace=FALSE)
```

Shuffle the order of an array

```{r}
sample(a, size=length(a), replace=FALSE)
```

Make sure you have enough elements

```{r, error=TRUE}
sample(1:5, size=10, replace=FALSE)
```

## Array constants

The `letters` and `LETTERS` constants hold lower and upper case letters

```{r}
letters[1:5]
LETTERS[1:5]
```

## Array operations

Many functions in R are *vectorized*, so they apply to arrays.

```{r}
a * 2
```

Compare individual elements

```{r}
a > 15
```

Compare each element across arrays

```{r}
a == c(10, 9, 12, 13, 14, 15, 16, 17, 18, 19, 20)
```

Select elements using boolean array

```{r}
a[a>15]
```

You can perform operations on the elements of two arrays **even if they are different sizes**. The smaller one wraps around.

```{r}
a = 1:5
b = rep(1, 8)
a + b
```

## Array functions

Length

```{r}
length(20:50)
```

Reverse

```{r}
rev(1:5)
```

Math

```{r}
sum(1:5)
min(1:5)
max(1:5)
```

`min()` and `max()` are not vectorized. They only return one value.
```{r}
max(1:5, 11:15, 21:25)
```

For a vectorized min and max, use `pmin()` and `pmax()`. P does not stand for "vectorized", but let's pretend it does.
```{r}
pmax(1:5, 11:15, 21:25)
```

## Array sorting

Sort

```{r}
a = c(70, 20, 80, 20, 10, 40)
sort(a)
```

Reverse

```{r}
sort(a, decreasing=TRUE)
```

Get the indices of the sorted values

```{r}
order(a)
```

## Test membership

To see if an item is in an array, use `%in%`

```{r}
9 %in% 1:10
9:11 %in% 1:10
```
