# Libraries and packages

A library or package is a collection of variables, datasets, functions, and/or operators.

It's called a "package" when being installed `install.packages("tidyverse")` and a "library" when being loaded for use `library(tidyverse)`.

A library and a package are the same thing, but R people will insist there is a difference. Whenever talking to R people, you've got a 50-50 chance of getting it right. If you get it wrong, you're going to get a short lecture. Just nod, and say "yes, that makes sense, and the distinction is clearly important". If you say anything else, you'll get a much longer more boring lecture.

I define these functions, so I don't have to worry about confusing the two.

```{r}
install.library = install.packages
package = library
```

If you only want to access one function or variable in a library without loading the whole thing, you can use `::`

```{r}
dplyr::band_instruments
```

## Errors when installing {#error_install}

Like a troll restricting access to a bridge, R will sometimes ask convoluted riddles when installing and especially when updating libraries.

### Restart?

> One or more of the packages to be updated are currently loaded. Restarting R prior to install is highly recommended.
>
> RStudio can restart R before installing the reuested packages. All work and data will be preserved during restart.
>
> Do you want to restart R prior to install?
>
> `[Yes]`\
> `[No]`\
> `[Cancel]` \<-- No one will blame you if you decide the complexity isn't worth it

The safer option is `[Yes]`, but you may end up in a ask-restart loop. So try `[Yes]`, but then click `[No]` if you're asked a second time.

You'll have to reload your libraries afterwards by redoing any `library(mylibrary)` calls you've already made.

### Locked directory?

> Error in install.packages : ERROR: failed to lock directory ...\
> Try removing '.../00LOCK'

Unfortunately, R is **not** compatible with any sort of software that backs up your documents folder. That includes Dropbox, Google Drive, and several others. To work around the problem, either:

1.  Delete the `00Lock` directory whenever R complains and retry.
2.  Close or turn off backups while using RStudio.

### Sources? Compilation?

> Do you want to install from sources the package which needs compilation?

Don't worry if you're confused. This convaluted question commonly stumps people. Here is a rephrasing:

> We don't have the latest version of this package for your system. We can (1) try to compile the latest version from source code or (2) install a slightly older version. What would you like to do?
>
> `[Yes]` = `[Try to compile and install new version]` This option is slow and can sometimes fail.\
> `[No]` = `[Install older version]` \<-- You usually want this option.\
> `[Cancel]` = `[Don't install]` Choose this if you don't think the libray is worth all this effort.
